---
title: "R Notebook"
output: html_notebook
---

Metrics:
Gini Coefficient is sometimes used in classification problems.
Gini coefficient can be straigh away derived from the AUC ROC number.
Gini is nothing but ratio between area between the ROC curve and the diagnol line 
            &
the area of the above triangle. Following is the formulae used :

Gini = 2*AUC – 1

Gini above 60% is a good model. 
Note:Not to be confused with the Gini index or Gini impurity, used when building decision trees.



example:
#normalizedGini(1:10000, runif(10000)) # Should be close to zero 
#normalizedGini(1:1000, 1:1000) 
#normalizedGini(1:1000, 1000:1) 



```{r}
normalizedGini <- function(aa, pp) {
    Gini <- function(a, p) {
        if (length(a) !=  length(p)) stop("Actual and Predicted need to be equal lengths!")
        temp.df <- data.frame(actual = a, pred = p, range=c(1:length(a)))
        temp.df <- temp.df[order(-temp.df$pred, temp.df$range),]
        population.delta <- 1 / length(a)
        total.losses <- sum(a)
        null.losses <- rep(population.delta, length(a)) # Hopefully is similar to accumulatedPopulationPercentageSum
        accum.losses <- temp.df$actual / total.losses # Hopefully is similar to accumulatedLossPercentageSum
        gini.sum <- cumsum(accum.losses - null.losses) # Not sure if this is having the same effect or not
        sum(gini.sum) / length(a)
    }
    Gini(aa,pp) / Gini(aa,aa)
}
```



```{r}
require(MLmetrics)
NormalizedGini()

```
