---
title: "R Notebook"
output: html_notebook
---

Steps Performed:

Understanding the data.
Finding missing values.

Separate categorical and numerical.
univariate analysis.
checking the distribution of each variable in numerical data.
checking the correlation for numerical data.
checking the frequency for categorical data.





Data contains 34 variables.

Household_ID ->	Unique identifier for each household.
Vehicle	-> Unique identifier for each vehicle within a household.
Calendar_Year ->	Calendar year during which vehicle was insured ,years:2005,2006,2007.
Model_Year ->	Model year of vehicle ,29levels->1981 to 2009
Blind_Make ->	Vehicle make(example: ACME, coded A)
Blind_Model ->	Vehicle model(example: ACME Roadster, coded A.1)
Blind_Submodel ->	Vehicle submodel (example: ACME Roadster LS, coded A.1.1) e.g. no suffix after the model name--submodel is coded as 0.

Categorical vehicle variable -> cat1 to cat12
Ordered categorical vehicle variable -> OrdCat 
Continuous vehicle variable, mean 0 stdev 1 -> Var1 to Var7
Categorical non-vehicle variable -> NVCat
Continuous non-vehicle variable, mean 0 stdev 1	-> NVVar1 to NVVar4



Target:Claim_Amount
The goal of this competition is to predict Bodily Injury Liability Insurance claim payments based on the characteristics of the insured’s vehicle.   

So it is a Regression Problem.



```{r}
Path="G:\\Old_drive\\Projects\\Allstate_Claim_Prediction_Challenge"
setwd(Path) #setting working directory
getwd()

data <- read.csv("data/train_1L.csv") #Reading the data
attach(data)     #attach the dataset ,we can call the variables directly without dataframe name.
names(data)   #column names of the data
```

```{r}
str(data) #structure of the data
```
checking the datatypes

```{r}
summary(data)
```



missing values(?)

Cat2 has 36897
Cat3 has 24
Cat4 has 42780
Cat5 has 42820
Cat6 has 213
Cat7 has 54386
Cat8 has 24
Cat10 has 26
Cat11 has 264
Cat12(blank) 241





Target variable:

```{r}
#Claim_Amount
#table(Claim_Amount)   #Claim_Amount frequency
prop.table(table(Claim_Amount)) #Claim_Amount frequency percentage.
```

99286 rows has 0's. which contains 99% data.



distribution of target variable
```{r}
# Density and rug  
d <- density(Claim_Amount)
plot(d, type="n", main="Claim_Amount")
polygon(d, col="lightgray", border="gray")
rug(Claim_Amount, col="red")
```

high concentrationof data is at 0.and gets sprased while the claim amount increases.



#plots
Distribution of categorical and numerical variables
```{r}
source(paste0(Path,"/scripts/utilities/plots_func.R"))  #reading plots function

plotsfunc(data,"G:\\Old_drive\\Projects\\Allstate_Claim_Prediction_Challenge\\plots") #plots will save into the path specified


```

Model-year=>left skewed.
var1 ,var2, var3 ,var4 ,var6 ,var7,var8 => right skewed.(long tail)

most the variables has outliers.



Seperate numerical and Categorical data

```{r}
factors <- sapply(data, is.factor) #checking the variable is a factor
vars_categ <- data[, factors]     #taking only factor variables
names(vars_categ)

vars_numeric <- data[,!factors]   #taking only numerical variables
names(vars_numeric)
```
17-categorical
18-Numerical



```{r}
length(unique(Household_ID)) #unique Household_ID
length(unique(Row_ID)) #unique Row_ID
```

Row id is unique.
but household id is dupicated.

```{r}
#Removing rowId
vars_numeric <- vars_numeric[, -which(names(vars_numeric) %in% c("Row_ID"))]

```


```{r}
length(Household_ID[duplicated(Household_ID)]) #duplicated records

```



```{r}
source(paste0(Path,"/scripts/utilities/correlation.R"))  #reading correlation
corFun(vars_numeric)
```





<!-- ```{r} -->
<!-- #pairs(vars_numeric) -->
<!-- cor(vars_numeric) -->
<!-- ``` -->

<!-- high correlated variables(0.74-0.99) -->

<!-- var1 & var3 = 0.743946820 -->
<!-- var1 & var5 = 0.905651716 -->
<!-- var1 & var6 = 0.777026876 -->
<!-- var2 & var4 = 0.982659211 -->
<!-- var2 & var6 = 0.770531671 -->
<!-- var3 & var5 = 0.784011528 -->
<!-- var3 & var6 = 0.825888541 -->
<!-- var4 & var6 = 0.784532509 -->
<!-- var5 & var6 = 0.749018242 -->
<!-- var6 & var7 = 0.80144159 -->






Saving Raw numeric and categorical data
```{r}

write.csv(vars_numeric,paste0(Path,'/data_for_models/raw_numerical.csv'),row.names = F)
write.csv(vars_categ,paste0(Path,'/data_for_models/raw_categorical.csv'),row.names = F)
          
```


```{r}
```





