---
title: "R Notebook"
output: html_notebook
---


Preprocessing steps:
loading categorical data.
rename the factor levels for Cat12.
created rowSums of missing values for cat1 to cat12.
checking the % of missing values in the data.
impute the missing values with mode imputation.
split the columns Blind_Model and Blind_subModel to get Vehicle details.
creaed dummy variables for cat1 to cat12.
find linear combos and remove it.
categorical encoding.
saving the file.





Reading the categorical data

```{r}
Path="G:\\Old_drive\\Projects\\Allstate_Claim_Prediction_Challenge"
vars_categ <- read.csv(paste0(dirname(getwd()),"/data_for_models/raw_categorical.csv")) #Reading the data
names(vars_categ)   #column names of the categorical data
```


```{r}
str(vars_categ)

```

rename the factor levels for Cat12 ->space with ?
```{r}
#vars_categ[vars_categ=="?"] 
table(vars_categ$Cat12)
levels(vars_categ$Cat12)
levels(vars_categ$Cat12)[levels(vars_categ$Cat12)==""] <- "?"
levels(vars_categ$Cat12)
table(vars_categ$Cat12)
```
  
  
Adding the missing values in row-wise for columns Cat 1 to 12 and made a new column with cat1_12_NAcount

```{r}

Cat1_12_NAcount <- data.frame(Cat1_12_NAcount=apply(vars_categ[,4:15],1,function(x) sum(x=='?')))
vars_categ <- cbind(vars_categ,Cat1_12_NAcount=Cat1_12_NAcount)

table(vars_categ$Cat1_12_NAcount)

```


again Adding the missing values in row-wise for columns Blind_Make,Blind_Model,Blind_Submodel.

```{r}
Blind1_3_NAcount <- data.frame(Blind1_3_NAcount=apply(vars_categ[,1:3],1,function(x) sum(x=='?')))
vars_categ <- cbind(vars_categ,Blind1_3_NAcount=Blind1_3_NAcount)

table(vars_categ$Blind1_3_NAcount)
```




% of missing values.
```{r}

source(paste0(Path,"/scripts/utilities/missvalpercent.R"))  #reading missing values percentages

missValpercent(vars_categ,"?")

```

Blind_Make ,Blind_Model ,Blind_Submodel has equal missing percentage.
Cat2 = 36%
Cat4 = 42%
Cat5 = 42%
Cat7 = 54%
and remaining has low missing values.



Imputing the missing categorical data 
```{r}

vars_categ[vars_categ=="?"] <- NA
apply(vars_categ,2,function(x) sum(is.na(x)))

sum(is.na(vars_categ))



source(paste0(Path,"/scripts/utilities/numeric(mean)-categorical(mode)Imputation.R"))  #reading plots function
vars_categ_imp <- Imputation(vars_categ)


#library(missForest)
#vars_categ_imp <- missForest(vars_categ[,c(4:16)], maxiter = 10, ntree = 100,verbose = TRUE,mtry = floor(sqrt(ncol(vars_categ[,c(4:16)]))))


sum(is.na(vars_categ_imp))
```


split the column Blind_Model into 2 parts
```{r}

 Blind_Model_split<- strsplit(as.character(vars_categ$Blind_Model),'.',fixed = T)
 Blind_Model_split<- do.call(rbind, Blind_Model_split)
 head(Blind_Model_split)

```


split the column Blind_Submodel into 3 parts
```{r}
 Blind_Submodel_split<- strsplit(as.character(vars_categ$Blind_Submodel),'.',fixed = T)
 Blind_Submodel_split<- do.call(rbind, Blind_Submodel_split)
 head(Blind_Submodel_split)
```

converting into dataframe and changing column names
```{r}
Blind_Submodel_split <- data.frame(Blind_Submodel_split)
names(Blind_Submodel_split) <- c('Vehicle_make','Vehicle_model','Vehicle_submodel')
head(Blind_Submodel_split)
```


```{r}
write.csv(Blind_Submodel_split,paste0(Path,'/data_for_models/vehical_details.csv'),row.names = F)
write.csv(vars_categ_imp,paste0(Path,'/data_for_models/vars_categ_imp.csv'),row.names = F)
```


#creating dummy for cat1 to cat12

```{r}
library(dummies)
vars_categ_dummy <- dummy.data.frame(vars_categ[,c(4:15)], sep = ".")

```

```{r}
write.csv(vars_categ_dummy,paste0(Path,'/data_for_models/vars_categ_dummy.csv'),row.names = F)
```

can be removed to eliminate the linear dependencies:
```{r}
library(caret)
Linear_comb <- findLinearCombos(vars_categ_dummy)
Linear_comb
```


```{r}
write.csv(vars_categ_dummy[,-Linear_comb$remove],paste0(Path,'/data_for_models/vars_categ_dummy_witoutlineardep.csv'),row.names = F)
```

categorical encoding 
```{r}
source(paste0(Path,"/scripts/utilities/Data_encoding.R"))
cat <- Data_encoding(vars_categ_imp[,4:15])
```

```{r}
write.csv(cat,paste0(Path,'/data_for_models/categ_encoding.csv'),row.names = F)
```
